---
title: "Untitled"
author: "Rachel Smith"
date: "5/28/2021"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

Create Unique data set:
```{r}
###install packages:
library(tidyverse)
library(readxl)
library(magrittr)
library(stringr)
library(lubridate)
library(here)
library(ggplot2)
library(tigerstats)

EEdata <- read_excel(here::here("/data/CleanedExtremeEventsData_2_27_20.xlsx"), sheet = 1)

###convert all spatial/temporal variables to factors
EEdata$Sampling_Duration_After <- factor(EEdata$Sampling_Duration_After)
EEdata$Sampling_Duration_Before <- factor(EEdata$Sampling_Duration_Before)
EEdata$Sampling_Duration_During <- factor(EEdata$Sampling_Duration_During)
EEdata$SamplingUnit_SpatialExtent <- factor(EEdata$SamplingUnit_SpatialExtent)
EEdata$Study_SpatialExtent <- factor(EEdata$Study_SpatialExtent)
EEdata$ProximateEvent_SpatialExtent <- factor(EEdata$ProximateEvent_SpatialExtent)
EEdata$ProximateEvent_Duration <- factor(EEdata$ProximateEvent_Duration)

###reduce EEdata to only unique accession numbers
EEunique <- {EEdata %>%
    distinct(UniqueAccession, .keep_all = T)
}

```


*Study Spatial Extent vs. Event Spatial Extent:*
```{r}

spatial<- xtabs(~ Study_SpatialExtent + ProximateEvent_SpatialExtent, data = EEunique)


```


*Study Spatial Extent vs. Event Spatial Extent, as bubble plot*
Here, bubbles are the number of effect sizes, not the proportion; heat maps in previous version of figure were proportions, not raw numbers, see "time_spatial_proportions.Rmd." 

```{r}
spatial_df <- as.data.frame(spatial)

spatial_df$Study_SpatialExtent <- factor(spatial_df$Study_SpatialExtent, levels = c("<1 sq m", "1-10 sq m", "10-100 sq m", "100-1000 sq m", ">1000 sq m", "Regional/Continental", "undefined"))

spatial_df$ProximateEvent_SpatialExtent <- factor(spatial_df$ProximateEvent_SpatialExtent, levels = c("<1 sq m", "1-10 sq m", "10-100 sq m", "100-1000 sq m", ">1000 sq m", "Regional/Continental"))


###Area of circle scaled to count
spatial.bub.plot<- ggplot(spatial_df, aes(x= Study_SpatialExtent, y =ProximateEvent_SpatialExtent, size = Freq)) + geom_point() + theme_bw() + 
  theme(axis.text.x = element_text(angle = 45, hjust = 1, color = "black", size = 14),  axis.title.x= element_text(size = 16),
        axis.title.y = element_text(size = 16), 
        plot.title = element_text(size = 22, hjust = .5, face = "bold"),
        strip.text = element_text(size=16), 
        legend.text=element_text(size=10),
        axis.text.y = element_text(color = "black", size = 14),
        panel.grid.major = element_blank(), panel.grid.minor = element_blank()) + 
  ylab("Event spatial extent") + xlab("Study spatial extent") +scale_size_area(breaks = c(0, 3, 6, 9, 12, 15))+ labs(size = "Number of responses")

# scale_size_area()

pdf("Spatial_Bubble_Plots.pdf", height = 6, width = 8)
spatial.bub.plot
dev.off()

```



*Sampling Unit Spatial Extent vs. Event Spatial Extent:*
  
  Here, rows are Sampling Unit Spatial Extent and columns are Event Spatial extent.

```{r sample}
sample<- xtabs(~ SamplingUnit_SpatialExtent + ProximateEvent_SpatialExtent, data = EEunique)
sample
```


*Sampling Unit Spatial Extent vs. Event Spatial Extent, as bubble plot*

```{r}
sample_df <- as.data.frame(sample)

sample_df$SamplingUnit_SpatialExtent <- factor(sample_df$SamplingUnit_SpatialExtent, levels = c("<1 sq m", "1-10 sq m", "10-100 sq m", "100-1000 sq m", ">1000 sq m", "Regional/Continental", "undefined"))


sample_df$ProximateEvent_SpatialExtent <- factor(sample_df$ProximateEvent_SpatialExtent, levels = c("<1 sq m", "1-10 sq m", "10-100 sq m", "100-1000 sq m", ">1000 sq m", "Regional/Continental"))


sample.bub.plot<- ggplot(sample_df, aes(x= SamplingUnit_SpatialExtent, y =ProximateEvent_SpatialExtent, size = Freq)) + geom_point() + theme_bw() + 
  theme(axis.text.x = element_text(angle = 45, hjust = 1),
        axis.text = element_text(color = "black")) + theme(axis.text = element_text(size = 12), axis.title.x= element_text(size = 16),
                                                           axis.title.y = element_text(size = 16), plot.title = element_text(size = 22, hjust = .5, face = "bold"),
                                                           strip.text = element_text(size=16), legend.text=element_text(size=10),
                                                           axis.text.x = element_text(angle = 45, hjust = 1, color = "black", size = 14),
                                                           axis.text.y = element_text(color = "black", size = 14),
                                                           panel.grid.major = element_blank(), panel.grid.minor = element_blank()) + ylab("Event spatial extent") + xlab("Sampling unit spatial extent") + scale_size_area(breaks = c(0, 2, 4, 6, 8, 10))+ labs(size = "Number of responses")


pdf("Sampling_Bubble_Plots.pdf", height = 6, width = 8)
sample.bub.plot
dev.off()
```


*Sampling Duration Before vs. Event Duration:*
  
  Here, rows are Sampling Duration Before and columns are Event Duration extent.

```{r before}

before1<- xtabs(~ Sampling_Duration_Before + ProximateEvent_Duration, data = EEunique)
before1

```


*Sampling Duration Before vs. Event Duration, as bubble plot:*
  
  Here, rows are Sampling Duration Before and columns are Event Duration extent.

```{r before}

before_df <- as.data.frame(before1)


before_df$Sampling_Duration_Before <- factor(before_df$Sampling_Duration_Before, levels = c("none", "single sample", "days", "weeks", "months", "years"))


before_df$ProximateEvent_Duration <- factor(before_df$ProximateEvent_Duration, levels = c("hours", "days", "weeks", "months", "years"))


before.bub.plot<- ggplot(before_df, aes(x= Sampling_Duration_Before, y =ProximateEvent_Duration, size = Freq)) + geom_point() + theme_bw() + 
  theme(axis.text.x = element_text(angle = 45, hjust = 1),
        axis.text = element_text(color = "black")) + theme(axis.text = element_text(size = 12), axis.title.x= element_text(size = 16),
                                                           axis.title.y = element_text(size = 16), plot.title = element_text(size = 22, hjust = .5, face = "bold"),
                                                           strip.text = element_text(size=16), legend.text=element_text(size=10),
                                                           axis.text.x = element_text(angle = 45, hjust = 1, color = "black", size = 14),
                                                           axis.text.y = element_text(color = "black", size = 14),
                                                           panel.grid.major = element_blank(), panel.grid.minor = element_blank()) + ylab("Event duration") + xlab("Sampling duration before event") +  scale_size_area(breaks = c(0, 3, 6, 9, 12, 15)) + labs(size = "Number of responses")


pdf("Before_Bubble_Plots.pdf", height = 6, width = 8)
before.bub.plot
dev.off()
```


*Sampling Duration During vs. Event Duration:*
  
Here, rows are Sampling Duration During and columns are Event Duration extent.

```{r during}


during1<- xtabs(~ Sampling_Duration_During + ProximateEvent_Duration, data = EEunique)
during1

```



*Sampling Duration During vs. Event Duration, as bubble plot:*
  
  Here, rows are Sampling Duration During and columns are Event Duration extent.

```{r during}

during_df <- as.data.frame(during1)



during.bub.plot<- ggplot(during_df, aes(x= Sampling_Duration_During, y =ProximateEvent_Duration, size = Freq)) + geom_point() + theme_bw() + 
  theme(axis.text.x = element_text(angle = 45, hjust = 1),
        axis.text = element_text(color = "black")) + theme(axis.text = element_text(size = 12), axis.title.x= element_text(size = 16),
                                                           axis.title.y = element_text(size = 16), plot.title = element_text(size = 22, hjust = .5, face = "bold"),
                                                           strip.text = element_text(size=16), legend.text=element_text(size=10),
                                                           axis.text.x = element_text(angle = 45, hjust = 1, color = "black", size = 14),
                                                           axis.text.y = element_text(color = "black", size = 14),
                                                           panel.grid.major = element_blank(), panel.grid.minor = element_blank()) + ylab("Event duration") + xlab("Sampling duration during event") +  scale_size_area(breaks = c(0, 3, 6, 9, 12, 15)) + labs(size = "Number of responses")


pdf("During_Bubble_Plots.pdf", height = 6, width = 8)
during.bub.plot
dev.off()

```


*Sampling Duration After vs. Event Duration:*
  
  Here, rows are Sampling Duration After and columns are Event Duration extent.

```{r after}

after1<- xtabs(~ Sampling_Duration_After + ProximateEvent_Duration, data = EEunique)
after1

```


*Sampling Duration After vs. Event Duration, as bubble plot:*
  
  Here, rows are Sampling Duration After and columns are Event Duration extent.

```{r after}
after_df <- as.data.frame(after1)



after.bub.plot<- ggplot(after_df, aes(x= Sampling_Duration_After, y =ProximateEvent_Duration, size = Freq)) + geom_point() + theme_bw() + 
  theme(axis.text.x = element_text(angle = 45, hjust = 1),
        axis.text = element_text(color = "black")) + theme(axis.text = element_text(size = 12), axis.title.x= element_text(size = 16),
                                                           axis.title.y = element_text(size = 16), plot.title = element_text(size = 22, hjust = .5, face = "bold"),
                                                           strip.text = element_text(size=16), legend.text=element_text(size=10),
                                                           axis.text.x = element_text(angle = 45, hjust = 1, color = "black", size = 14),
                                                           axis.text.y = element_text(color = "black", size = 14),
                                                           panel.grid.major = element_blank(), panel.grid.minor = element_blank()) + ylab("Event duration") + xlab("Sampling duration after event") +  scale_size_area(breaks = c(0, 2, 4, 6, 8, 10)) + labs(size = "Number of responses") 


pdf("After_Bubble_Plots.pdf", height = 6, width = 8)
after.bub.plot
dev.off()

```
