---
  title: "Untitled"
author: "Rachel Smith & Brandon Sansom"
date: "Update on 12/04/2019"
output: github_document
---
  
  ```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, echo = F, message = F, error = F, warning = F)
```

### Exploratory plots of space/time variables: 

```{r prepare}
###install packages:
library(tidyverse)
library(readxl)
library(magrittr)
library(stringr)
library(lubridate)
library(here)


EEdata <- read_excel(here::here("CleanedExtremeEventsData_8Nov2019.xlsx"), sheet = 1)

###convert all spatial/temporal variables to factors

EEdata$Sampling_Duration_After <- factor(EEdata$Sampling_Duration_After)
EEdata$Sampling_Duration_Before <- factor(EEdata$Sampling_Duration_Before)
EEdata$Sampling_Duration_During <- factor(EEdata$Sampling_Duration_During)
EEdata$SamplingUnit_SpatialExtent <- factor(EEdata$SamplingUnit_SpatialExtent)
EEdata$Study_SpatialExtent <- factor(EEdata$Study_SpatialExtent)
EEdata$ProximateEvent_SpatialExtent <- factor(EEdata$ProximateEvent_SpatialExtent)
EEdata$ProximateEvent_Duration <- factor(EEdata$ProximateEvent_Duration)




###look at factor levels:
#unique(EEdata$Sampling_Duration_After) ##years, months, none, weeks, single sample, days
#unique(EEdata$Sampling_Duration_Before)##years, months, none, weeks, single sample, days
#unique(EEdata$Sampling_Duration_During) #years, months, none, weeks, single sample, days, hours
#unique(EEdata$SamplingUnit_SpatialExtent)##<1 sqm, >1000 sq m, 1-10 sq m, 10-100 sq m, 100-1000 sq m, Regional/Continent, undefined
#unique(EEdata$Study_SpatialExtent)#<1 sqm, >1000 sq m, 1-10 sq m, 10-100 sq m, 100-1000 sq m, Regional/Continent, undefined
#unique(EEdata$ProximateEvent_Duration)###days, hours, months, weeks, years; this is missing single sample and none
#unique(EEdata$ProximateEvent_SpatialExtent)###<1 sqm, >1000 sqm, 10-100 sqm, 100-1000 sqm, Regional/Continent; this one is missing 1-10 sqm

```

#### Contingency tables:

*Study Spatial Extent vs. Event Spatial Extent:*
  
  Here, rows are Study Spatial Extent and columns are Event Spatial extent.
```{r spatial}

###https://www.datacamp.com/community/tutorials/contingency-tables-r
####here the rows are Study Spatial extent and and columns are Event Spatial extent
###We can also turn these into proportions/do a chi square test

spatial<- xtabs(~ Study_SpatialExtent + ProximateEvent_SpatialExtent, data = EEdata)
spatial

#table(EEdata$Study_SpatialExtent, EEdata$ProximateEvent_SpatialExtent)

###rows and columns are not independent
#chisq.test(EEdata$Study_SpatialExtent, EEdata$ProximateEvent_SpatialExtent)
```

*Sampling Unit Spatial Extent vs. Event Spatial Extent:*
  
  Here, rows are Sampling Unit Spatial Extent and columns are Event Spatial extent.

```{r sample}
sample<- xtabs(~ SamplingUnit_SpatialExtent + ProximateEvent_SpatialExtent, data = EEdata)
sample


#table(EEdata$SamplingUnit_SpatialExtent, EEdata$ProximateEvent_SpatialExtent)

###rows and columns are not independent
#chisq.test(EEdata$SamplingUnit_SpatialExtent, EEdata$ProximateEvent_SpatialExtent)
```

*Sampling Duration Before vs. Event Duration:*
  
  Here, rows are Sampling Duration Before and columns are Event Duration extent.

```{r before}

before1<- xtabs(~ Sampling_Duration_Before + ProximateEvent_Duration, data = EEdata)
before1

#before<- table(EEdata$Sampling_Duration_Before, EEdata$ProximateEvent_Duration)
#before
###rows and columns are not independent
#chisq.test(EEdata$Sampling_Duration_Before, EEdata$ProximateEvent_Duration)
```


*Sampling Duration During vs. Event Duration:*
  
  Here, rows are Sampling Duration During and columns are Event Duration extent.

```{r during}


during1<- xtabs(~ Sampling_Duration_During + ProximateEvent_Duration, data = EEdata)
during1

#during<- table(EEdata$Sampling_Duration_During, EEdata$ProximateEvent_Duration)
#during
###rows and columns are not independent
#chisq.test(EEdata$Sampling_Duration_During, EEdata$ProximateEvent_Duration)

```


*Sampling Duration After vs. Event Duration:*
  
  Here, rows are Sampling Duration After and columns are Event Duration extent.

```{r after}

after1<- xtabs(~ Sampling_Duration_After + ProximateEvent_Duration, data = EEdata)
after1

#after<-table(EEdata$Sampling_Duration_After, EEdata$ProximateEvent_Duration)
#after
###rows and columns are not independent
#chisq.test(EEdata$Sampling_Duration_After, EEdata$ProximateEvent_Duration)
```

#### Heat Maps: 

library(ggplot2)

*Study Spatial Extent vs. Event Spatial Extent:*
  ```{r sample1,  fig.width = 12, fig.height = 12}
spatial_df <- as.data.frame(spatial)

spatial_df$Study_SpatialExtent = factor(spatial_df$Study_SpatialExtent,levels(spatial_df$Study_SpatialExtent)[c(7,1,3:5,2,6)])
spatial_df$ProximateEvent_SpatialExtent = factor(spatial_df$ProximateEvent_SpatialExtent,levels(spatial_df$ProximateEvent_SpatialExtent)[c(1,3,4,2,5)])

ggplot(spatial_df, aes(Study_SpatialExtent, ProximateEvent_SpatialExtent)) +
  geom_tile(aes(fill = Freq), colour = "black") +
  scale_fill_gradient(low = "white", high = "steelblue")
```

*Sampling Unit Spatial Extent vs. Event Spatial Extent:*
  ```{r sample1,  fig.width = 12, fig.height = 12}
sample_df <- as.data.frame(sample)

sample_df$SamplingUnit_SpatialExtent = factor(sample_df$SamplingUnit_SpatialExtent,levels(sample_df$SamplingUnit_SpatialExtent)[c(7,1,3:5,2,6)])
sample_df$ProximateEvent_SpatialExtent = factor(sample_df$ProximateEvent_SpatialExtent,levels(sample_df$ProximateEvent_SpatialExtent)[c(1,3,4,2,5)])

ggplot(sample_df, aes(SamplingUnit_SpatialExtent, ProximateEvent_SpatialExtent)) +
  geom_tile(aes(fill = Freq), colour = "black") +
  scale_fill_gradient(low = "white", high = "steelblue")
```


